---
title: "The Ideal Electoral System"
author: "Daniel K Baissa"
date: "3/28/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(stargazer)
library(foreign)

Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jdk-17.0.1')

#Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jdk-11.0.1') #laptop
require(rJava)
options(java.parameters = "-Xmx75g")     # or 8g, or larger than this, ...
require(bartMachine)
set_bart_machine_num_cores(30)

df <- read.dta("Data/datar.dta")
```

# Electoral Engineering

Today we have access to machine learning models that can help use greatly improve our ability to develop electoral systems.

# Carey and Hix Replication 

I will start by replicating a model from the Carry and Hix dataset in order to ensure that the data and models line up before improving the specification.

### The first model from Carey and Hix 2011

Here I will replicate the coefficients for their model and omit the Panel Corrected Standard Errors

```{r}

CH <- df %>% 
  select(disprop, dist_mag_medians, dm_asym, pres, legal_thresh, MMPL, compensatory, 
         ethnic_fract_fearon, hybrid , election_yr, pol_freedom, econ_freedom, 
         population, gdp_head, growth , age_dem, federal, latitude, col_uk, 
         col_sp_po, col_oth, americas, former_com, pacific, s_asia,  gini, africa_me)

m.lm <- lm(disprop ~ ., data = CH)
summary(m.lm)


```

### Enter BART

Now let's run Carey and Hix's model using Bayesian Additive Regression Trees. The key difference is that we are going to remove their asymptotic term from the model since BART will automatically allow for curvilinear fits. Then we will rename the y variable "y".

```{r}
X <- df %>% 
  select(disprop, 
         dist_mag_medians, pres, legal_thresh, MMPL, compensatory, 
         ethnic_fract_fearon, hybrid , election_yr, pol_freedom, econ_freedom, 
         population, gdp_head, growth , age_dem, federal, latitude, col_uk, 
         col_sp_po, col_oth, americas, former_com, pacific, s_asia,  gini, africa_me) %>% 
  na.omit()

y <- X$disprop

X <- X %>% 
  select(!disprop)
```

```{r}
m.BART <- bartMachine(X = X, y = y)
m.BART
```

We can see here that the Pseudo R Squared for the BART model is significantly higher than that of their OLS model. 

```{r}
plot_convergence_diagnostics(m.BART)

```

```{r}

plot_y_vs_yhat(m.BART)

```

```{r}
pd_plot(m.BART,  m.BART[["training_data_features"]][1])
```



Great! The main effect mirrors the researcher's findings very well, but with much wider confidence intervals.

# Optimal Governance 

Carey and Hix focused used their models to estimate which district magnitudes would be best for optomizing governance. This means for example, they took the median of the `disprop` variable and then said values below the median were idea and those above where less than ideal. 

Let's try something a bit more varied. Let's try spliting the data into 3rds and then take the bottom third as ideal.

```{r}
X <- df %>% 
  select(disprop, 
         dist_mag_medians, pres, legal_thresh, MMPL, compensatory, 
         ethnic_fract_fearon, hybrid , election_yr, pol_freedom, econ_freedom, 
         population, gdp_head, growth , age_dem, federal, latitude, col_uk, 
         col_sp_po, col_oth, americas, former_com, pacific, s_asia,  gini, africa_me) %>%
  mutate(ideal_disprop = ifelse(disprop < quantile(disprop, probs = seq(0, 1, by = 1/3))[[2]], 1, 0)) %>% 
  na.omit()

y <- as.factor(X$ideal_disprop)

X <- X %>% 
  select(!c(disprop, ideal_disprop))

```



```{r}
m.BART <- bartMachine(X = X, y = y)
m.BART
```

```{r}
pd_plot(m.BART,  m.BART[["training_data_features"]][1])
```


Here we see that a high district magnitude will be needed holding all else equal.


```{r}

y_hat_class = predict(m.BART, X, type = "class")


test <- cbind(df, "Predicted Ideal" = y_hat_class) %>% 
  select(country, `Predicted Ideal`, election_yr, growth, gini, federal, pres) 

```

## Accountability

```{r}
X <- df %>% 
  select(enps, 
         dist_mag_medians, pres, legal_thresh, MMPL, compensatory, 
         ethnic_fract_fearon, hybrid , election_yr, pol_freedom, econ_freedom, 
         population, gdp_head, growth , age_dem, federal, latitude, col_uk, 
         col_sp_po, col_oth, americas, former_com, pacific, s_asia,  gini, africa_me) %>%
  mutate(ideal_enps = ifelse(enps < quantile(enps, probs = seq(0, 1, by = 1/3))[[2]], 1, 0)) %>% 
  na.omit()

y <- as.factor(X$ideal_enps)

X <- X %>% 
  select(!c(enps, ideal_enps))
```


```{r}
m.BART <- bartMachine(X = X, y = y)
m.BART
```

```{r}
pd_plot(m.BART,  m.BART[["training_data_features"]][1])
```

# Are there "Ideal" countries?


Let's say we want better than median Accountability and better than 2/3rds of representation
```{r}
X <- df %>% 
  select(enps, disprop, country,
         dist_mag_medians, pres, legal_thresh, MMPL, compensatory, 
         ethnic_fract_fearon, hybrid , election_yr, pol_freedom, econ_freedom, 
         population, gdp_head, growth , age_dem, federal, latitude, col_uk, 
         col_sp_po, col_oth, americas, former_com, pacific, s_asia,  gini, africa_me) %>%
  mutate(ideal_enps = ifelse(enps < quantile(enps, probs = seq(0, 1, by = 1/2))[[2]], 1, 0)) %>% 
  mutate(ideal_disprop = ifelse(enps < quantile(disprop, probs = seq(0, 1, by = 1/3))[[2]], 1, 0)) %>% 
  # select(country, ideal_enps, ideal_disprop) %>% 
  mutate(ideal = ifelse(ideal_enps == 1 & ideal_disprop == 1, 1, 0)) %>% 
  na.omit()

y <- as.factor(X$ideal)

X <- X %>% 
  select(!c(enps, ideal_enps, ideal_disprop, ideal, country, disprop))
```

```{r}
m.BART <- bartMachine(X = X, y = y)
m.BART
```

```{r}
pd_plot(m.BART,  m.BART[["training_data_features"]][1])
```

Here we can see that District Magnitude is Significant and there is no reason to limit the threshold to an arbitrarily low number of seats.

# Can we improve a country?

Let's take Chile, can we improve Chile?

```{r}
Chile <- df %>% 
  select(enps, disprop, country,
         dist_mag_medians, pres, legal_thresh, MMPL, compensatory, 
         ethnic_fract_fearon, hybrid , election_yr, pol_freedom, econ_freedom, 
         population, gdp_head, growth , age_dem, federal, latitude, col_uk, 
         col_sp_po, col_oth, americas, former_com, pacific, s_asia,  gini, africa_me) %>%
  mutate(ideal_enps = ifelse(enps < quantile(enps, probs = seq(0, 1, by = 1/2))[[2]], 1, 0)) %>% 
  mutate(ideal_disprop = ifelse(enps < quantile(disprop, probs = seq(0, 1, by = 1/3))[[2]], 1, 0)) %>% 
  # select(country, ideal_enps, ideal_disprop) %>% 
  mutate(ideal = ifelse(ideal_enps == 1 & ideal_disprop == 1, 1, 0)) %>% 
  filter(country == "Chile")

```
Let's confirm that Chile is not seen as an ideal country.

```{r}

Chile_changed = predict(m.BART, X, type = "class")

Chile_changed <- cbind(Chile_changed, X)

Chile_changed %>% 
  filter(ethnic_fract_fearon < 0.498) %>% 
  filter(ethnic_fract_fearon > 0.496) 
```

It's not.

Now let's see what would happen if we changed their electoral rules to be more representative.

```{r}

X2 <- X %>% 
  mutate(dist_mag_medians = ifelse(ethnic_fract_fearon < 0.498 & ethnic_fract_fearon > 0.496,
                                   20 , dist_mag_medians))
```

```{r}
Chile_changed = predict(m.BART, X2, type = "class")

Chile_changed <- cbind(Chile_changed, X2)

Chile_changed %>% 
  filter(ethnic_fract_fearon < 0.498) %>% 
  filter(ethnic_fract_fearon > 0.496) 

```

Its not enough...


Let's also add federlaism?

```{r}
X2 <- X %>% 
  # mutate(dist_mag_medians = ifelse(ethnic_fract_fearon < 0.498 & ethnic_fract_fearon > 0.496,
  #                                  6 , dist_mag_medians))
  # # %>% 
  mutate(federal = ifelse(ethnic_fract_fearon < 0.498 & ethnic_fract_fearon > 0.496,
                                   1 , federal))

Chile_changed = predict(m.BART, X2, type = "class")

Chile_changed <- cbind(Chile_changed, X2)

Chile_changed %>% 
  filter(ethnic_fract_fearon < 0.498) %>% 
  filter(ethnic_fract_fearon > 0.496) 
  
```

Federalism seems to fix it?!?! Federalism is more important than magnitude.

Since that is the case, let's look at if Federalism is statistically significant.

```{r}
pd_plot(m.BART,  m.BART[["training_data_features"]][15])
```

Interesting! On average Federalism makes things worse. In Chile, however, Federalism makes things better. This means that the BART model is finding tailored electoral and governmental rules for each country.

# Variable selection

Perhaps its possible that there are more important predictors of ideal governance.

```{r}
var_sel <- var_selection_by_permute(m.BART,
                                    bottom_margin = 2)
```

